---
title: "Employee Retention"
author: "Willie McRae"
date: "25 July 2016"
output: html_document
---
# Data Science Meetup - Imperial Museum 25th July 
This was my approach during the meetup. I focused mainly on exploratory views.

# Outline
Running through Exercise 3 from the excellent [A collection of Data Science Take-home Challenges](http://datascientistjobinterview.com/)

1. Assume, for each company, that the headcount starts from zero on 2011/01/23. 
2. Estimate employee headcount, for each company, on each day, from 2011/01/24 to 2015/12/13.
That is, if by 2012/03/02 2000 people have joined company 1 and 1000 of them have
already quit, then company headcount on 2012/03/02 for company 1 would be 1000.
3. You should create a table with 3 columns: day, employee_headcount, company_id.
4. What are the main factors that drive employee churn? Do they make sense? Explain your
findings.
5. If you could add to this data set just one variable that could help explain employee churn,
what would that be?


# Load
``` {r load, warning = FALSE, message = FALSE}
# Set directory
setwd("/Users/wbidst/Desktop/R/Employee_Retention")

# Load packages
install.packages("plyr", repos="http://cran.rstudio.com/")
library(plyr)
install.packages("dplyr", repos="http://cran.rstudio.com/")
library(dplyr)
install.packages("stringr", repos="http://cran.rstudio.com/")
library(stringr)
install.packages("tidyr", repos="http://cran.rstudio.com/")
library(tidyr)
install.packages("ggplot2", repos="http://cran.rstudio.com/")
library(ggplot2)
install.packages("lubridate", repos="http://cran.rstudio.com/")
library(lubridate)
install.packages("rattle", repos="http://cran.rstudio.com/")
library(rattle) 
install.packages("rpart.plot", repos="http://cran.rstudio.com/")
library(rpart.plot)
install.packages("RColorBrewer",repos="http://cran.rstudio.com/")
library(RColorBrewer) 
library(rpart) 

# Load data
d <- read.csv("employee_retention_data.csv", header = TRUE)

```


# Clean
``` {r clean}
# Check structure
str(d)
summary(d)

# Add a variable for Exited 'Status' - H = Hardcore, still employed Q = Quitter
d$status <- ifelse(d$quit_date == "NA", "H", "Q")
d$status[which(is.na(d$status))] <- "H"
d$status <- as.factor(d$status)
str(d)


# Correct class of columns with dates
d$join_date <- as.Date(d$join_date)
d$quit_date <- as.Date(d$quit_date) 


# Add a variable called tenure
# First change the NA values to a new date, later than the latest date
d2 <- d # Keeping d as my backup!
d2$quit_date[which(is.na(d$quit_date))] <- "2016-07-25"
d2$tenure <- d2$quit_date - d2$join_date
d2$tenure <- as.numeric(d2$tenure) 

summary(d2)
# Quitters and Hardcore 

quitters <- filter(d2, status == "Q")
summary(quitters)
hardcore <- filter(d2, status == "H")
summary(hardcore)
```


# Tidy
``` {r tidy}
# All rows are observations, all columns are variables - nothing to do here??
# Did not get to actually create the table suggested in exercise
```


# Explore 
``` {r explore}
require(scales)

# Basic visualizations
## Histograms
### Salary
ggplot(data = d, aes(x = salary)) +
  geom_histogram() +
  facet_wrap(~ company_id, nrow = 4) +
  scale_x_continuous(labels = comma)

ggplot(data = d, aes(x = salary, fill = status)) +
  geom_histogram(position = "fill") +
  facet_wrap(~ company_id, nrow = 4) +
  scale_x_continuous(labels = comma) # People paid a lot don't exit??

### Seniority
ggplot(data = d, aes(x = seniority)) +
  geom_histogram() +
  facet_wrap(~ company_id, nrow = 4) +
  scale_x_continuous(labels = comma)

ggplot(data = d, aes(x = seniority, fill = status)) +
  geom_histogram(position = "fill") +
  facet_wrap(~ company_id, nrow = 4) +
  scale_x_continuous(labels = comma) 

### Tenure
ggplot(data = d2, aes(x = tenure)) +
  geom_histogram() +
  facet_wrap(~ company_id, nrow = 4) +
  scale_x_continuous(labels = comma)

ggplot(data = d2, aes(x = tenure, fill = status)) +
  geom_histogram(position = "fill") +
  facet_wrap(~ company_id, nrow = 4) +
  scale_x_continuous(labels = comma) 

# Scatter plots
ggplot(data = d, aes(x = seniority, y = salary, col = status)) +
  geom_point(shape = 21, alpha = 0.5, position = "jitter") +
  scale_y_continuous(labels = comma) +
  facet_grid(.~ status) 

ggplot(data = d, aes(x = factor(company_id), y = salary, col = status)) +
  geom_point(alpha = 0.5, position = "jitter") +
  scale_y_continuous(labels = comma) +
  facet_grid(.~ status) # There are a couple of outliers in company 1 and 10 (super senior people?)

ggplot(data = d, aes(x = seniority, y = salary, col = status)) +
  geom_point(alpha = 0.5, position = "jitter") +
  scale_y_continuous() +
  facet_grid(company_id ~ status)

# Bar plots
ggplot(data = d, aes(x = company_id, y = salary, fill = status, group = status)) +
  geom_bar(stat = "identity", position = "stack") 


# Box plots
ggplot(data = d2, aes(x = factor(company_id), y = tenure, fill = status)) +
  geom_boxplot() 

ggplot(data = d2, aes(x = factor(company_id), y = salary, fill = status)) +
  geom_boxplot() 

ggplot(data = d2, aes(x = factor(dept), y = tenure, fill = status)) +
  geom_boxplot() 

ggplot(data = d2, aes(x = factor(dept), y = salary, fill = status)) +
  geom_boxplot() 

# A quick look at tenure
ggplot(data = d2, aes(x = join_date, y = quit_date, fill = status)) +
  geom_point(shape = 21, alpha = 0.5) +
  geom_smooth()

ggplot(data = filter(d2, status == "Q"), aes(x = factor(company_id), y = tenure, fill = status)) +
  geom_boxplot() 

```


# Estimate headcount per company per day 
``` {r estimate_peoplecount}
# Create table that shows headcount flux by company

# Start with a vector of every day
Days <- seq(as.Date("2011-01-24"), as.Date("2015-12-13"), by = "days")

# Now, create a blank dataframe
daydf <- data.frame(Date = Days)

# Try ifelse statements to populate (only need company_id and employee_headcount)
# Gave up on this did not prioritize

```


# Identify key variables in churn 
``` {r identify_variables1}
# Which variables are most important in determining exit?

tree <- rpart(status ~ company_id +seniority + salary + tenure + dept, data = d2, method = "class")
fancyRpartPlot(tree) # Tenure appears to be the best predictive variable. 

tree2 <- rpart(status ~ company_id + seniority + salary + dept, data = d2, method = "class")
fancyRpartPlot(tree2) # High salaries stick around


```


# Add new variable 
``` {r identify_variables}
# What new variable might help explain churn
# I think the tenure variable added does this nicely
# Would like to see 'date last promoted' or something similar

```


# Build predictive model 
``` {r build}
# Which variables are most important in determining exit? Apply these to the remaining set.
# Would like to apply characteristics of Quitters to Hardcore - did not get to this.

```
